Opening Restaurants During COVID-19 Can Prove Beneficial for Restaurant Revenue in the GTA

### 2021-02-27 outputs/Paper.Rmd
update

# Introduction
Current COVID-19 lockdown policies are having a significant effect on restaurant businesses within the GTA. Previous research shows that over half of Canadian restaurants may be unable to operate in current conditions that were set out by the local government. This can be detrimental to the economic situation as restaurant businesses can contribute over 4% of the national GDP. As a result, the government needs to consider the business effects of these lockdown policies on restaurants across the GTA. A randomized controlled trial was conducted to investigate these factors. Within the treatment group, restaurants were given an intervention that allowed them to open for indoor/patio dining while a control group remained open only for delivery and take-out services. After one month, these restaurants were given a survey that asked them to indicate different business-related factors, such as their expense changes, employee salaries, revenue, and COVID-related exposures during the experimental condition. More details with regards to data collection, intervention, descriptive analysis, and overview of findings will be provided in subsequent sections. Finally, limitations of the study as well as future directions will be discussed. This analysis will ultimately inform decisions with regards to the severity of restrictions imposed on restaurants in the GTA during the current pandemic.

@@

## Overview 
Previous research has demonstrated that 60% of Canadian restaurants may be unable to operate with current policies in response to the Coronavirus within three months. Moreover, 1.2 million Canadians are said to be employed within the restaurant industry with an annual payroll of $30 billion. Restaurants contribute a significant amount to Canada’s economy, and particularly in the GTA. However, it is still unknown how much restaurants receive through business during the pandemic. According to Perrin Beatty (CEO of the Canadian Chamber of Commerce), restaurants may find it “nearly impossible” to make money during COVID-19 lockdown restrictions. Moreover, ongoing COVID restrictions contribute to the employment rate of the restaurant industry to fall by 21% below pre-COVID levels. It was predicted that half of table-service restaurants expect to reduce their employed staff due to strict lockdown policies. By researching the effects of allowing restaurants to offer dine-in services, we can better understand the economic impact of re-opening restaurants to the public. It is also important to consider the risk associated with re-opening restaurants as well. According to a recent model by provincial health officials, the number of restaurant-related outbreaks of COVID appear significantly lower than the number of school, child-care, or long-term care related outbreaks. Dr. Colin Furness of the University of Toronto cited that this number is an underrepresentation of restaurant related transmissions, due to a lack of contact tracing. As is the case, measurements of COVID-related incidents within restaurants that offer dine-in services versus restaurants that don’t may provide clarity on this matter. Overall, in order to decide on appropriate lockdown policies within the restaurant industry, further clarity is required to make an informed decision. As existing contributions rely heavily on predictive speculation, it is important to verify their claims in the most recent light (February 2021). Furthermore, contributions seem to apply to national restaurants across Canada more broadly. By applying a randomized control trial to the GTA, researchers are able to distinguish nuance at a regional level. This experimental method can then be applied to other areas across Canada in order to establish more appropriate government policies region by region.
## Findings

>>>>

# Introduction
Current COVID-19 lockdown policies are having a significant effect on restaurant businesses within the GTA. Previous research shows that over half of Canadian restaurants may be unable to operate in current conditions that were set out by the local government [@toronto_2020]. This can be detrimental to the economic situation as restaurant businesses can contribute over 4% of the national GDP [@lew_2020]. As a result, the government needs to consider the business effects of these lockdown policies on restaurants across the GTA. A randomized controlled trial was conducted to investigate these factors. Within the treatment group, restaurants were given an intervention that allowed them to open for indoor/patio dining while a control group remained open only for delivery and take-out services. After one month, these restaurants were given a survey that asked them to indicate different business-related factors, such as their expense changes, employee salaries, revenue, and COVID-related exposures during the experimental condition. More details with regards to data collection, intervention, descriptive analysis, and overview of findings will be provided in subsequent sections. Finally, limitations of the study as well as future directions will be discussed. This analysis will ultimately inform decisions with regards to the severity of restrictions imposed on restaurants in the GTA during the current pandemic.

@@

## Overview 
Previous research has demonstrated that 60% of Canadian restaurants may be unable to operate with current policies in response to the Coronavirus within three months. Moreover, 1.2 million Canadians are said to be employed within the restaurant industry with an annual payroll of $30 billion [@toronto_2020]. Restaurants contribute a significant amount to Canada’s economy, and particularly in the GTA. However, it is still unknown how much restaurants receive through business during the pandemic. According to Perrin Beatty (CEO of the Canadian Chamber of Commerce), restaurants may find it “nearly impossible” to make money during COVID-19 lockdown restrictions [@toronto_2020]. Moreover, ongoing COVID restrictions contribute to the employment rate of the restaurant industry to fall by 21% below pre-COVID levels. It was predicted that half of table-service restaurants expect to reduce their employed staff due to strict lockdown policies [@restaurants_canada]. By researching the effects of allowing restaurants to offer dine-in services, we can better understand the economic impact of re-opening restaurants to the public. It is also important to consider the risk associated with re-opening restaurants as well. According to a recent model by provincial health officials, the number of restaurant-related outbreaks of COVID appear significantly lower than the number of school, child-care, or long-term care related outbreaks. Dr. Colin Furness of the University of Toronto cited that this number is an underrepresentation of restaurant related transmissions, due to a lack of contact tracing [@declerq_2020]. As is the case, measurements of COVID-related incidents within restaurants that offer dine-in services versus restaurants that don’t may provide clarity on this matter. Overall, in order to decide on appropriate lockdown policies within the restaurant industry, further clarity is required to make an informed decision. As existing contributions rely heavily on predictive speculation, it is important to verify their claims in the most recent light (February 2021). Furthermore, contributions seem to apply to national restaurants across Canada more broadly. By applying a randomized control trial to the GTA, researchers are able to distinguish nuance at a regional level. This experimental method can then be applied to other areas across Canada in order to establish more appropriate government policies region by region.
## Findings
